---
title: "Analysis for story and graphics: Takeaways from 2020 census for Maryland, in 5 charts"
output: html_document
author: Aadit Tambe
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction
This data notebook contains the analysis that generated facts in the story. For each sentence in the story generated by original data analysis, we have provided the corresponding code and results.


## Load libraries

```{r}

library(tidyverse)
library(janitor)

```
## Load data
Capital News Service examined 2020 census data released by the U.S. Census Bureau. 

The data set, titled "Hispanic or Latino, and not Hispanic or Latino by Race" with Table ID: P2, can be found in  [here](https://data.census.gov/cedsci/table?q=United%20States&g=0400000US24&y=2020&d=DEC%20Redistricting%20Data%20%28PL%2094-171%29&tid=DECENNIALPL2020.P2).

```{r}
data_md_2020 <- read_csv("data/DECENNIALPL2020.P2_data_with_overlays_2021-09-30T125006.csv", 
    skip = 1) %>% clean_names()

data_md_2010 <- read_csv("data/DECENNIALPL2010.P2_data_with_overlays_2021-09-30T162358.csv", 
    skip = 1) %>% clean_names()

data_md_counties_2020<- read_csv("data/DECENNIALPL2020.P2_data_with_overlays_2021-09-30T161838.csv", 
    skip = 1) %>% clean_names()

data_md_counties_2010<- read_csv("data/DECENNIALPL2010.P2_data_with_overlays_2021-09-30T162053.csv", 
    skip = 1) %>% clean_names()


data_us_counties_2010<- read_csv("data/DECENNIALPL2010.P2_data_with_overlays_2021-09-30T210749.csv", skip = 1) %>% 
  clean_names()

data_us_counties_2020<- read_csv("data/DECENNIALPL2020.P2_data_with_overlays_2021-09-30T210749.csv", skip = 1) %>% 
  clean_names()

md_above_18_2020<- read.csv("data/DECENNIALPL2020.P4_data_with_overlays_2021-09-30T225914.csv", skip = 1) %>% 
  clean_names()
```
### Analysis for graphics

#### Graphic 1: Maryland map, with population by race (Maryland map)
```{r}
# Analysis for graphic showing Maryland map, with population by race

counties_2020 = data_md_counties_2020 %>% 
   select(geographic_area_name, total, total_hispanic_or_latino, total_not_hispanic_or_latino_population_of_one_race_white_alone, total_not_hispanic_or_latino_population_of_one_race_black_or_african_american_alone, total_not_hispanic_or_latino_population_of_one_race_american_indian_and_alaska_native_alone, total_not_hispanic_or_latino_population_of_one_race_asian_alone, total_not_hispanic_or_latino_population_of_one_race_native_hawaiian_and_other_pacific_islander_alone, total_not_hispanic_or_latino_population_of_one_race_some_other_race_alone, total_not_hispanic_or_latino_population_of_two_or_more_races) %>% 
  rename(hispanic = total_hispanic_or_latino, white = total_not_hispanic_or_latino_population_of_one_race_white_alone, black = total_not_hispanic_or_latino_population_of_one_race_black_or_african_american_alone, native = total_not_hispanic_or_latino_population_of_one_race_american_indian_and_alaska_native_alone, asian = total_not_hispanic_or_latino_population_of_one_race_asian_alone, pacific_islander = total_not_hispanic_or_latino_population_of_one_race_native_hawaiian_and_other_pacific_islander_alone, other = total_not_hispanic_or_latino_population_of_one_race_some_other_race_alone, two_races = total_not_hispanic_or_latino_population_of_two_or_more_races) %>% 
  pivot_longer(-geographic_area_name, names_to = "races", values_to = "count") %>% 
  mutate(year = 2020) %>% 
#filter(races != "total", geographic_area_name != "United States")
filter(geographic_area_name != "United States")


counties_2010 = data_md_counties_2010 %>% 
   select(geographic_area_name, total, total_hispanic_or_latino, total_not_hispanic_or_latino_population_of_one_race_white_alone, total_not_hispanic_or_latino_population_of_one_race_black_or_african_american_alone, total_not_hispanic_or_latino_population_of_one_race_american_indian_and_alaska_native_alone, total_not_hispanic_or_latino_population_of_one_race_asian_alone, total_not_hispanic_or_latino_population_of_one_race_native_hawaiian_and_other_pacific_islander_alone, total_not_hispanic_or_latino_population_of_one_race_some_other_race_alone, total_not_hispanic_or_latino_two_or_more_races) %>% 
  rename (hispanic = total_hispanic_or_latino, white = total_not_hispanic_or_latino_population_of_one_race_white_alone, black = total_not_hispanic_or_latino_population_of_one_race_black_or_african_american_alone, native = total_not_hispanic_or_latino_population_of_one_race_american_indian_and_alaska_native_alone, asian = total_not_hispanic_or_latino_population_of_one_race_asian_alone, pacific_islander = total_not_hispanic_or_latino_population_of_one_race_native_hawaiian_and_other_pacific_islander_alone, other = total_not_hispanic_or_latino_population_of_one_race_some_other_race_alone, two_races = total_not_hispanic_or_latino_two_or_more_races) %>% 
  mutate(total = as.numeric(total)) %>% 
  pivot_longer(-geographic_area_name, names_to = "races", values_to = "count") %>% 
  mutate(year = 2010) 
  #filter( races != "total")

counties_2010_2020 <- rbind(counties_2010, counties_2020)

summary_1 = counties_2010_2020 %>% 
  pivot_wider(names_from = year, values_from = count) %>% 
  rename(year_2020 = `2020`, year_2010 = `2010`)

# this includes totals
md_counties_graphic_final = summary_1 %>% 
  mutate(pct_change = (((year_2020 - year_2010)/year_2010)*100)) %>% 
  select(-year_2020, -year_2010) %>% 
  pivot_wider(names_from = races, values_from = pct_change)

#t <- md_counties_graphic_final %>% 
 # pivot_wider(names_from = cou, values_from = pct_change)

# this is ready to be joined with flourish json
write.csv(md_counties_graphic_final,"results/md_counties_graphic_final.csv", row.names = FALSE)


```


#### Graphic 2: Percent population share by race (pie chart)
```{r}
# Analysis for md graphic


md_2020 = data_md_2020 %>% 
   select(geographic_area_name, total, total_hispanic_or_latino, total_not_hispanic_or_latino_population_of_one_race_white_alone, total_not_hispanic_or_latino_population_of_one_race_black_or_african_american_alone, total_not_hispanic_or_latino_population_of_one_race_american_indian_and_alaska_native_alone, total_not_hispanic_or_latino_population_of_one_race_asian_alone, total_not_hispanic_or_latino_population_of_one_race_native_hawaiian_and_other_pacific_islander_alone, total_not_hispanic_or_latino_population_of_one_race_some_other_race_alone, total_not_hispanic_or_latino_population_of_two_or_more_races) %>% 
  rename(hispanic = total_hispanic_or_latino, white = total_not_hispanic_or_latino_population_of_one_race_white_alone, black = total_not_hispanic_or_latino_population_of_one_race_black_or_african_american_alone, native = total_not_hispanic_or_latino_population_of_one_race_american_indian_and_alaska_native_alone, asian = total_not_hispanic_or_latino_population_of_one_race_asian_alone, pacific_islander = total_not_hispanic_or_latino_population_of_one_race_native_hawaiian_and_other_pacific_islander_alone, other = total_not_hispanic_or_latino_population_of_one_race_some_other_race_alone, two_races = total_not_hispanic_or_latino_population_of_two_or_more_races) %>% 
  pivot_longer(-geographic_area_name, names_to = "races", values_to = "count") %>% 
  mutate(year = 2020) %>% 
filter(races != "total", geographic_area_name != "United States")



md_2010 = data_md_2010 %>% 
   select(geographic_area_name, total, total_hispanic_or_latino, total_not_hispanic_or_latino_population_of_one_race_white_alone, total_not_hispanic_or_latino_population_of_one_race_black_or_african_american_alone, total_not_hispanic_or_latino_population_of_one_race_american_indian_and_alaska_native_alone, total_not_hispanic_or_latino_population_of_one_race_asian_alone, total_not_hispanic_or_latino_population_of_one_race_native_hawaiian_and_other_pacific_islander_alone, total_not_hispanic_or_latino_population_of_one_race_some_other_race_alone, total_not_hispanic_or_latino_two_or_more_races) %>% 
  rename (hispanic = total_hispanic_or_latino, white = total_not_hispanic_or_latino_population_of_one_race_white_alone, black = total_not_hispanic_or_latino_population_of_one_race_black_or_african_american_alone, native = total_not_hispanic_or_latino_population_of_one_race_american_indian_and_alaska_native_alone, asian = total_not_hispanic_or_latino_population_of_one_race_asian_alone, pacific_islander = total_not_hispanic_or_latino_population_of_one_race_native_hawaiian_and_other_pacific_islander_alone, other = total_not_hispanic_or_latino_population_of_one_race_some_other_race_alone, two_races = total_not_hispanic_or_latino_two_or_more_races) %>% 
  mutate(total = as.numeric(total)) %>% 
  pivot_longer(-geographic_area_name, names_to = "races", values_to = "count") %>% 
  mutate(year = 2010) %>% 
  filter( races != "total")

md_2010_2020 <- rbind(md_2010, md_2020)

summary_2 = md_2010_2020 %>% 
  pivot_wider(names_from = year, values_from = count)%>% 
  rename(year_2020 = `2020`, year_2010 = `2010`) %>% 
  mutate(percent_in_2010 = ((year_2010)/sum(year_2010))*100) %>% 
  mutate(percent_in_2020 = ((year_2020)/sum(year_2020))*100) 

md_state_graphic_final = summary_2 %>% 
  mutate(pct_change = (((year_2020  - year_2010)/year_2010)*100)) %>% 
  select(-year_2020, -year_2010) %>% 
  pivot_wider(names_from = races, values_from = pct_change)

# analysis for md race pie chart

#we have md_state_graphic_final
sum(summary_2$'year_2010')
sum(summary_2$'year_2020')

md_race_data <- summary_2 %>% 
  # mutate(pct_in_2010 = ((summary_2$'year_2010')/sum(summary_2$'year_2010'))*100) %>% 
  # mutate(pct_in_2020 = ((summary_2$'year_2020')/sum(summary_2$'year_2020'))*100) 
  mutate(pct_change = ((year_2020-year_2010)/year_2010)*100)


sum(md_race_data$pct_in_2010)


```

#### Graphic 3: Counties around Washington, D.C. with population by race (D.C. counties map)
```{r}
# analysis for the dc graphic

target <- c ("Arlington County, Virginia", "Charles County, Maryland","Alexandria city, Virginia","Stafford County, Virginia","Spotsylvania County, Virginia","Calvert County, Maryland","Fauquier County, Virginia","Jefferson County, West Virginia","Culpeper County, Virginia","Manassas city, Virginia","Warren County, Virginia","Fredericksburg city, Virginia","Fairfax city, Virginia","Fairfax County, Virginia","Manassas Park city, Virginia","Clarke County, Virginia","Falls Church city, Virginia","Madison County, Virginia","Rappahannock County, Virginia", "Frederick County, Maryland", "Loudoun County, Virginia", "Montgomery County, Maryland", "Prince George's County, Maryland", "Prince William County, Virginia", "District of Columbia, District of Columbia")

filtered_dc_counties_2020 = data_us_counties_2020 %>% 
   select(geographic_area_name, total, total_hispanic_or_latino, total_not_hispanic_or_latino_population_of_one_race_white_alone, total_not_hispanic_or_latino_population_of_one_race_black_or_african_american_alone, total_not_hispanic_or_latino_population_of_one_race_american_indian_and_alaska_native_alone, total_not_hispanic_or_latino_population_of_one_race_asian_alone, total_not_hispanic_or_latino_population_of_one_race_native_hawaiian_and_other_pacific_islander_alone, total_not_hispanic_or_latino_population_of_one_race_some_other_race_alone, total_not_hispanic_or_latino_population_of_two_or_more_races) %>% 
  rename(hispanic = total_hispanic_or_latino, white = total_not_hispanic_or_latino_population_of_one_race_white_alone, black = total_not_hispanic_or_latino_population_of_one_race_black_or_african_american_alone, native = total_not_hispanic_or_latino_population_of_one_race_american_indian_and_alaska_native_alone, asian = total_not_hispanic_or_latino_population_of_one_race_asian_alone, pacific_islander = total_not_hispanic_or_latino_population_of_one_race_native_hawaiian_and_other_pacific_islander_alone, other = total_not_hispanic_or_latino_population_of_one_race_some_other_race_alone, two_races = total_not_hispanic_or_latino_population_of_two_or_more_races) %>% 
  pivot_longer(-geographic_area_name, names_to = "races", values_to = "count") %>% 
  mutate(year = 2020) %>% 
#filter(races != "total", geographic_area_name != "United States")
filter(geographic_area_name != "United States") 

filtered_dc_counties_2020 = filter(filtered_dc_counties_2020, geographic_area_name %in% target)


filtered_dc_counties_2010 = data_us_counties_2010 %>% 
   select(geographic_area_name, total, total_hispanic_or_latino, total_not_hispanic_or_latino_population_of_one_race_white_alone, total_not_hispanic_or_latino_population_of_one_race_black_or_african_american_alone, total_not_hispanic_or_latino_population_of_one_race_american_indian_and_alaska_native_alone, total_not_hispanic_or_latino_population_of_one_race_asian_alone, total_not_hispanic_or_latino_population_of_one_race_native_hawaiian_and_other_pacific_islander_alone, total_not_hispanic_or_latino_population_of_one_race_some_other_race_alone, total_not_hispanic_or_latino_two_or_more_races) %>% 
  rename (hispanic = total_hispanic_or_latino, white = total_not_hispanic_or_latino_population_of_one_race_white_alone, black = total_not_hispanic_or_latino_population_of_one_race_black_or_african_american_alone, native = total_not_hispanic_or_latino_population_of_one_race_american_indian_and_alaska_native_alone, asian = total_not_hispanic_or_latino_population_of_one_race_asian_alone, pacific_islander = total_not_hispanic_or_latino_population_of_one_race_native_hawaiian_and_other_pacific_islander_alone, other = total_not_hispanic_or_latino_population_of_one_race_some_other_race_alone, two_races = total_not_hispanic_or_latino_two_or_more_races) %>% 
  mutate(total = as.numeric(total)) %>% 
  pivot_longer(-geographic_area_name, names_to = "races", values_to = "count") %>% 
  mutate(year = 2010) 
  #filter( races != "total")
  filtered_dc_counties_2010 = filter(filtered_dc_counties_2010, geographic_area_name %in% target)


dc_2010_2020 <- rbind(filtered_dc_counties_2010, filtered_dc_counties_2020)

summary_3 = dc_2010_2020 %>% 
  pivot_wider(names_from = year, values_from = count)%>% 
  rename(year_2020 = `2020`, year_2010 = `2010`)

# this includes totals
dc_counties_graphic_final = summary_3 %>% 
  mutate(pct_change = (((year_2020  - year_2010)/year_2010)*100)) %>% 
  select(-year_2020, -year_2010) %>% 
  pivot_wider(names_from = races, values_from = pct_change)

# this is ready to be joined with flourish json
write.csv(dc_counties_graphic_final,"results/dc_counties_graphic_final.csv", row.names = FALSE)
```

#### Graphic 4: Maryland population 18 years of age and older, and under 18
```{r}
# graphic for md age

# analysis for over 18 data
filtered_md_above_18_2020<-md_above_18_2020 %>% 
   select(geographic_area_name, x_total, x_total_hispanic_or_latino, x_total_not_hispanic_or_latino_population_of_one_race_white_alone, x_total_not_hispanic_or_latino_population_of_one_race_black_or_african_american_alone, x_total_not_hispanic_or_latino_population_of_one_race_american_indian_and_alaska_native_alone, x_total_not_hispanic_or_latino_population_of_one_race_asian_alone, x_total_not_hispanic_or_latino_population_of_one_race_native_hawaiian_and_other_pacific_islander_alone, x_total_not_hispanic_or_latino_population_of_one_race_some_other_race_alone, x_total_not_hispanic_or_latino_population_of_two_or_more_races) %>% 
  rename(total = x_total, hispanic = x_total_hispanic_or_latino, white = x_total_not_hispanic_or_latino_population_of_one_race_white_alone, black = x_total_not_hispanic_or_latino_population_of_one_race_black_or_african_american_alone, native = x_total_not_hispanic_or_latino_population_of_one_race_american_indian_and_alaska_native_alone, asian = x_total_not_hispanic_or_latino_population_of_one_race_asian_alone, pacific_islander = x_total_not_hispanic_or_latino_population_of_one_race_native_hawaiian_and_other_pacific_islander_alone, other = x_total_not_hispanic_or_latino_population_of_one_race_some_other_race_alone, two_races = x_total_not_hispanic_or_latino_population_of_two_or_more_races) %>% 
  pivot_longer(-geographic_area_name, names_to = "races", values_to = "count") %>% 
  mutate(age = "over_18") %>% 
filter(races != "total", geographic_area_name != "United States")
#filter(geographic_area_name != "United States") 


#total md population 2020
md_age_total = data_md_2020 %>% 
   select(geographic_area_name, total, total_hispanic_or_latino, total_not_hispanic_or_latino_population_of_one_race_white_alone, total_not_hispanic_or_latino_population_of_one_race_black_or_african_american_alone, total_not_hispanic_or_latino_population_of_one_race_american_indian_and_alaska_native_alone, total_not_hispanic_or_latino_population_of_one_race_asian_alone, total_not_hispanic_or_latino_population_of_one_race_native_hawaiian_and_other_pacific_islander_alone, total_not_hispanic_or_latino_population_of_one_race_some_other_race_alone, total_not_hispanic_or_latino_population_of_two_or_more_races) %>% 
  rename(hispanic = total_hispanic_or_latino, white = total_not_hispanic_or_latino_population_of_one_race_white_alone, black = total_not_hispanic_or_latino_population_of_one_race_black_or_african_american_alone, native = total_not_hispanic_or_latino_population_of_one_race_american_indian_and_alaska_native_alone, asian = total_not_hispanic_or_latino_population_of_one_race_asian_alone, pacific_islander = total_not_hispanic_or_latino_population_of_one_race_native_hawaiian_and_other_pacific_islander_alone, other = total_not_hispanic_or_latino_population_of_one_race_some_other_race_alone, two_races = total_not_hispanic_or_latino_population_of_two_or_more_races) %>% 
  pivot_longer(-geographic_area_name, names_to = "races", values_to = "count") %>% 
  mutate(age = "all") %>% 
filter(races != "total", geographic_area_name != "United States")
# sum(md_age_total$count)


md_age_total_and_over_18 <- rbind(md_age_total, filtered_md_above_18_2020)

summary_4 = md_age_total_and_over_18 %>% 
  pivot_wider(names_from = age, values_from = count)


# now subtract over18 from total to get final result
md_age_graphic_final = summary_4 %>% 
  mutate(under_18 = (summary_4$'all' - summary_4$'over_18'))

md_age_graphic_final = md_age_graphic_final %>% 
  mutate(pct_over_18 = (md_age_graphic_final$'over_18' /(sum(md_age_graphic_final$'over_18')))*100)

# now get percentages
md_age_graphic_final = md_age_graphic_final %>% 
  mutate(pct_under_18 = (md_age_graphic_final$'under_18' /(sum(md_age_graphic_final$'under_18')))*100)
#sum(md_age_graphic_final$pct_under_18) # sum returns 100 whihc is good

# write.csv(md_age_graphic_final,"results/md_age_graphic_final.csv", row.names = FALSE)

```


# Graphic 5: Demographic change in counties around Baltimore
```{r}
# analysis for baltimore county graphics

# define counties in baltimore area
target <- c ("Baltimore city, Maryland", "Baltimore County, Maryland", "Harford County, Maryland", "Howard County, Maryland", "Carroll County, Maryland", "Anne Arundel County, Maryland")

# we already have md_counties_graphic_final
md_counties_graphic_final

counties_around_baltimore = filter(md_counties_graphic_final, geographic_area_name %in% target)
counties_around_baltimore

```


### Sentence analysis

#### Fact 1: "Maryland’s population increased by 7 percent to 6,177,224 in the last decade."

```{r}
total_pop_2010 <-  sum(summary_2$year_2010)
total_pop_2020 <-  sum(summary_2$year_2020)

# calculate percent increase
fact1 = ((total_pop_2020-total_pop_2010)/total_pop_2010)*100

# 6.991744
# Population increased by 7 percent.
```

#### Fact 2: "Census data show that there was a 10.5 percent decrease in the last decade in people who identify as white alone."
```{r}
fact2 <- summary_2 %>% 
  filter(races == "white")%>% 
  select(races, year_2010, year_2020) %>% 
  mutate(percent_change = ((year_2020-year_2010)/year_2010)*100)
# -7.732085
# White population decreased by 7.7 percent.
```

#### Fact 3: "Now, the white plurality makes up 48.7 percent of the state's total population."
```{r}
fact3 <- summary_2 %>% 
  filter(races == "white")%>% 
  select(races, percent_in_2020) 

#White population accounted for 47.16976% of the total population of the state in 2020.

```

#### Fact 4: "Black people, who make up 29 percent of Maryland’s population, are the second largest demographic group."
```{r}
fact4 <- summary_2 %>% 
  filter(races == "black")%>% 
  select(races, percent_in_2020) 

# Black population accounted for 29.0588% of the state's population in 2020.

```

#### Fact 5.1: "ll but three counties in Maryland — Frederick County, Worcester County and Queen Anne’s County — saw a decrease in white population while seeing an increase in Hispanic Marylanders..."
```{r}
# Decrease in white population
fact5_1a <- summary_1 %>% 
  filter(races == "white")%>% 
  select(geographic_area_name, year_2010, year_2020) %>% 
  mutate(pct_change = ((year_2020-year_2010)/year_2010)*100 )

# All but three counties in Maryland — Frederick County, Worcester County and Queen Anne’s County — saw a decrease in white population

# Increase in Hispanic population
fact5_1b <- summary_1 %>% 
  filter(races == "hispanic")%>% 
  select(geographic_area_name, year_2010, year_2020) %>% 
  mutate(pct_change = ((year_2020-year_2010)/year_2010)*100 )
# All counties saw an increase in Hispanic population


```


#### Fact 5.2: "... who make up 11.8% of the total population."
```{r}
fact5_2 <- summary_2 %>% 
  filter(races == "hispanic")%>% 
  select(races, percent_in_2020) 

# 	Hispanic population accoutned for 11.81348% of the state's population in 2020.
```



#### Fact 6: People who identify as “other races” increased nearly 194 percent, from 11,972 in 2010, to 35,314 in 2020 — the largest percent increase for a single demographic group. 
```{r}
fact6 <- summary_2 %>% 
  select(races, year_2010, year_2020) %>% 
  mutate(pct_change = ((year_2020-year_2010)/year_2010)*100 )

# Other races increased by 194 percent from 11,972 in 2010 to 35,314 in 2020. It was the largest percent increase for a single demographic group.
  
```


#### Fact 7: Marylanders of two or more races increased by 115 percent from 2010 to 2020. 
```{r}
fact7 <- summary_2 %>% 
  select(races, year_2010, year_2020) %>% 
  mutate(pct_change = ((year_2020-year_2010)/year_2010)*100 ) %>% 
  filter(races == "two_races")

# There was a 115.1653% increase in people identifying as two or more races
```

#### Fact 8: People under age 18 make up 22 percent of the state’s population. 
```{r}
under18 = sum(md_age_graphic_final$under_18)
#1362022

over18 = sum(md_age_graphic_final$over_18)
# 4815202

total = over18 + under18
#  6177224

fact8 = (under18/total)*100
# 22.04 percent are people under age 18

```


#### Fact 9: "Among that group, less than 37.6 percent of people are white, while a little under half are white among people 18 and over."
```{r}
fact9_1<- md_age_graphic_final

# 37.62215295 are white among people under 18

fact9_2<- md_age_graphic_final
# 49.87038965 are white among people over 18

```


#### Fact 10: "Hispanic people make up 17.4 percent of the population under age 18, while 10.2 percent of people are Hispanic among people 18 and over."

```{r}
fact10_1<- md_age_graphic_final
# 17.43606197 percent of population under age 18 is Hispanic

fact10_2<-md_age_graphic_final
# 10.22308098 percent of population over 19 is Hispanic
```


#### Fact 11: "Baltimore city’s population saw a 6 percent decrease, from 620,961 in 2010 to 585,708 in 2020." 
```{r}
baltimore<- summary_1 %>% 
  filter(geographic_area_name == "Baltimore city, Maryland", races == "total")
# Population in 2010: 620961
# Population in 2020: 585708
fact11 = ((585708-620961)/585708)*100
# Baltimore city’s population saw a 6 percent decrease, from 620,961 in 2010 to 585,708 in 2020.

```


#### Fact 12: "The counties surrounding Baltimore — Baltimore County, Anne Arundel County, Howard County, Harford County and Carroll County — saw an increase in population."
```{r}
fact12 = counties_around_baltimore %>% 
  select(geographic_area_name, total)

# All counties other than Baltimore saw an increase in population.

```

#### Fact 13: "Baltimore County, which wraps around the Baltimore city, saw 6.1 percent increase in population."
```{r}
fact13 <- counties_around_baltimore %>% 
  select(geographic_area_name, total) %>% 
  filter(geographic_area_name=="Baltimore County, Maryland")

# There was a 6.1 percent in Baltimore County's population.
```

#### Fact 14: "Maryland counties surrounding Washington, D.C. saw growth in population, with Montgomery County being the state’s largest county, with its population accounting for X percent of all Marylanders."
```{r}
fact14_1<-summary_1 %>% 
  filter(geographic_area_name == "Montgomery County, Maryland" |  geographic_area_name == "Prince George's County, Maryland") %>% 
  filter(races == "total")


fact14_2<-summary_1 %>% 
  filter(races == "total")


montgomery = 1062061
md_population = total_pop_2020

# Montgomery county is the most populous in the state.

montgomery_pct = (montgomery/md_population)*100

# Montgomery county accounts for 17.19318% of the state's population.
```




